---
title: "CPU analysis - Point to point arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'point_to_point', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements
```

# Data exploration
```{r}
measurements %>%
  group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_cpu, type = "full")

measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(avg_cpu, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "avg_cpu", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('CPU Utilisation (%)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "cpu_point_to_point_histogram.pdf")
```

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "avg_cpu", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('CPU Utilisation (%)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "cpu_point_to_point_box.pdf")
```
# ANOVA

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(avg_cpu ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey") + custom_theme
ggdensity(res, fill="lightgrey") + custom_theme

# Create a QQ plot of residuals
ggqqplot(res) + custom_theme
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal. Trying some transformations (sqrt, log, reciprocal) also gives no normal result.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(avg_cpu)
shapiro_tests[order(shapiro_tests$p),]
```

26 out of 32 groups have a p-value above 0.05, so we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "avg_cpu", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "avg_cpu", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "avg_cpu", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "avg_cpu", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, some points exhibit a large distance to the reference line The normality assumption is therefore not met.

## 4. Homogeneity of variance

```{r}
levene_test(avg_cpu ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is higher than 0.05, hence we fail to reject H0. Thus, we can assume homogeneity of variance  among all groups.

It is important to note that some groups have an extremely high variance (exceeding 1.5x the other variances).

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_cpu, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
The assumptions for ANOVA are violated, hence we resort to PERMANOVA (using permutations).

## Running PERMANOVA
```{r}
set.seed(2051)
aovp.out <- aovp(avg_cpu ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

The results indicate a significant main effect of the algorithm, map resolution and angular update factors.

There is also a significant interaction between the algorithm ad angular update.

### Post hoc tests
Since there are only two treatments for the map resolution (high and low) we do not need to perform a post hoc test for this main effect.

To analyse the main effect of the algorithm and angular update (interaction), we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'avg_cpu', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm, angular_update)  %>% shapiro_test(avg_cpu)
shapiro_tests[order(shapiro_tests$p),]
```

```{r}
ggqqplot(measurements, "avg_cpu", ggtheme = theme_bw()) + facet_grid(angular_update ~ algorithm, labeller = "label_both")
```

```{r}
levene_test(avg_cpu ~ algorithm : angular_update, data = measurements)
```

```{r}
agg_measurements <- measurements %>% group_by(algorithm, angular_update) %>%
  get_summary_stats(avg_cpu, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
The normality of groups assumption is violated, but homogeneity of variance cannot be rejected (p > 0.05).

The assumptions for Tukey's HSB test are not met, even after transformation. Hence, we resort to the non-parametric Dunn test.

```{r}
algorithm_angular_update_interaction <- interaction(measurements$algorithm, measurements$angular_update)
DunnTest(avg_cpu ~ algorithm_angular_update_interaction, measurements, method='BH')
```
