---
title: "ROS message analysis - Circular arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'point_to_point', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements$rosbag_msg_size <- measurements$rosbag_msg_size / 1048576
measurements
```

# Data exploration - Count
```{r}
measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(rosbag_msg_count, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "rosbag_msg_count", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('ROS message count') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist70
ggexport(hist, filename = "ros_message_count_point_to_point_histogram.pdf")
```

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "rosbag_msg_count", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('ROS message count') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "ros_message_count_point_to_point_box.pdf")
```
# ANOVA - Count

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(rosbag_msg_count ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey")
ggdensity(res, fill="lightgrey")

# Create a QQ plot of residuals
ggqqplot(res)
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal due to outliers. Trying some transformations gives no normal result either.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(rosbag_msg_count)
shapiro_tests[order(shapiro_tests$p),]
```

25 out of 32 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "rosbag_msg_count", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "rosbag_msg_count", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "rosbag_msg_count", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "rosbag_msg_count", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```

### 4. Homogeneity of variance
#### Transformed data
```{r}
levene_test(rosbag_msg_count ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(rosbag_msg_count, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```

## Running (PERM)ANOVA

```{r}
set.seed(2111)
aovp.out <- aovp(rosbag_msg_count ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

### Post hoc tests
To analyse the difference between algorithms, we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'rosbag_msg_count', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm)  %>% shapiro_test(rosbag_msg_count)
shapiro_tests[order(shapiro_tests$p),]
```

```{r}
measurements %>% ggqqplot("rosbag_msg_count", ggtheme = theme_bw(), fill="algorithm", color="algorithm")
```

```{r}
levene_test(rosbag_msg_count ~ algorithm, data = measurements)
```
The assumptions for Tukey's HSB test are not met. Hence, we resort to the non-parametric Dunn test.

```{r}
DunnTest(rosbag_msg_count ~ algorithm, measurements, method='BH')
```

#############################################################################
#############################################################################
#############################################################################
#############################################################################
#############################################################################
#############################################################################
#############################################################################
#############################################################################


# Data exploration - Size

```{r}
measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(rosbag_msg_size, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "rosbag_msg_size", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('ROS message size (MB)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "ros_message_size_circular_histogram.pdf")
```

To get an impression of the distribution of the distribution of average cpu utilisation per group, we plot the data in boxplots.

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "rosbag_msg_size", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('ROS message size (MB)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "ros_message_size_circular_box.pdf")
```
# ANOVA - Size
## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(rosbag_msg_size ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey")
ggdensity(res, fill="lightgrey")

# Create a QQ plot of residuals
ggqqplot(res)
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal due to outliers. Trying some transformations found that reciprocal transform gives no normal result either.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(rosbag_msg_size)
shapiro_tests[order(shapiro_tests$p),]
```

25 out of 32 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "rosbag_msg_size", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "rosbag_msg_size", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "rosbag_msg_size", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "rosbag_msg_size", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```
In the plots above, most points are around the normal line.  The normality assumption is therefore met.

### 4. Homogeneity of variance
#### Transformed data
```{r}
levene_test(rosbag_msg_size ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is lower than 0.05, hence we reject H0. Thus, we cannot assume homogeneity of variance among all groups.

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(rosbag_msg_size, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```

## Running (PERM)ANOVA

```{r}
set.seed(2121)
aovp.out <- aovp(rosbag_msg_size ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

### Post hoc test
To analyse the difference between algorithms, we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'rosbag_msg_size', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm)  %>% shapiro_test(rosbag_msg_size)
shapiro_tests[order(shapiro_tests$p),]
```

```{r}
measurements %>% ggqqplot("rosbag_msg_size", ggtheme = theme_bw(), fill="algorithm", color="algorithm")
```

```{r}
levene_test(rosbag_msg_size ~ algorithm, data = measurements)
```
The assumptions for Tukey's HSB test are not met. Hence, we resort to the non-parametric Dunn test.

```{r}
DunnTest(rosbag_msg_size ~ algorithm, measurements, method='BH')
```
