---
title: "Memory analysis - Circular arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'circular', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements$avg_mem <- measurements$avg_mem / 1048576
measurements
```

# Data exploration
```{r}
measurements %>%
  group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_mem, type = "full")

measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(avg_mem, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "avg_mem", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('Memory utilisation (MB)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "mem_circular_histogram.pdf")
```

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "avg_mem", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('Memory utilisation (MB)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "mem_circular_box.pdf")
```
# ANOVA

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(avg_mem ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey")
ggdensity(res, fill="lightgrey")

# Create a QQ plot of residuals
ggqqplot(res)
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal due to outliers.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(avg_mem)
shapiro_tests[order(shapiro_tests$p),]
```

28 out of 32 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "avg_mem", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```
In the plots above, most points are around the normal line. The normality assumption is therefore met. There do appear to be some outliers.

### 4. Homogeneity of variance
```{r}
levene_test(avg_mem ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is higher than 0.05, hence we fail to reject H0. Thus, we can assume homogeneity of variance among all groups.

When plotting the data we observe one group with an extremely high variance.

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_mem, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```

## Running (PERM)ANOVA
```{r}
set.seed(2061)
aovp.out <- aovp(avg_mem ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

The results indicate significant main effect of the algorithm and angular update factors.

There is also a significant interaction between the algorithm and map resolution.

### Post hoc tests
The angular update factor only has 2 treatments (high andl low), thus we do not have to perform any post hoc test for this effect.

#### Algorithm / Map resolution
To analyse the effect of the algorithm and map resolution interaction, we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'avg_mem', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm, map_resolution)  %>% shapiro_test(avg_mem)
shapiro_tests[order(shapiro_tests$p),]
```
5 out of 8 are not significant, indicating normality for these groups.

```{r}
ggqqplot(measurements, "avg_mem", ggtheme = theme_bw()) + facet_grid(map_resolution ~ algorithm, labeller = "label_both")
```
Especially for gmapping, the resulting plots seem non-normal.

```{r}
levene_test(avg_mem ~ algorithm : map_resolution, data = measurements)
```

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution) %>%
  get_summary_stats(avg_mem, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
One outlier (corresponding to gmapping) is apparent.
The assumptions for Tukey's HSB test are not met. Hence, we resort to the non-parametric Dunn test.

```{r}
algorithm_map_resolution_interaction <- interaction(measurements$algorithm, measurements$map_resolution)
DunnTest(avg_mem ~ algorithm_map_resolution_interaction, measurements, method='BH')
```
