---
title: "Memory analysis - Point to point arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'point_to_point', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements$avg_mem <- measurements$avg_mem / 1048576
measurements
```

# Data exploration
```{r}
measurements %>%
  group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_mem, type = "full")

measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(avg_mem, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "avg_mem", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('Memory utilisation (MB)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "mem_point_to_point_histogram.pdf")
```

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "avg_mem", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('Memory utilisation (MB)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "mem_point_to_point_box.pdf")
```

# ANOVA

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(avg_mem ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey") + custom_theme
ggdensity(res, fill="lightgrey") + custom_theme

# Create a QQ plot of residuals
ggqqplot(res) + custom_theme
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal. Trying some transformations (sqrt, log, reciprocal) also gives no normal result.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(avg_mem)
shapiro_tests[order(shapiro_tests$p),]
```

29 out of 32 groups have a p-value above 0.05, so we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "avg_mem", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "avg_mem", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, some points exhibit a large distance to the reference line. The normality assumption is however still met, since most points lie around the reference line (supported by the shapiro tests too).

## 4. Homogeneity of variance

```{r}
levene_test(avg_mem ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is lower than 0.05, hence we reject H0. Thus, we cannot assume homogeneity of variance among all groups.

This is likely due to some gmapping groups with an extremely high variance.

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(avg_mem, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
The assumptions for ANOVA are violated, hence we resort to PERMANOVA (using permutations).

## Running PERMANOVA
```{r}
set.seed(2071)
aovp.out <- aovp(avg_mem ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

The results indicate a significant main effect of the algorithm and map resolution factors.

There is also a significant interaction between the algorithm and map resolution.

### Post hoc tests
To analyse the main effect of the algorithm and map resolution (interaction), we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'avg_mem', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm, map_resolution)  %>% shapiro_test(avg_mem)
shapiro_tests[order(shapiro_tests$p),]
```
For half of the groups we can reject H0 (normality) because p < 0.05. The data does not seem normal.

```{r}
ggqqplot(measurements, "avg_mem", ggtheme = theme_bw()) + facet_grid(map_resolution ~ algorithm, labeller = "label_both")
```
The plots support our previous result.

```{r}
levene_test(avg_mem ~ algorithm : map_resolution, data = measurements)
```
When plotting the data we observe one group with an extremely high variance.

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution) %>%
  get_summary_stats(avg_mem, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```

The assumptions for Tukey's HSB test are not met, even after transformation. Hence, we resort to the non-parametric Dunn test.

```{r}
algorithm_map_resolution_interaction <- interaction(measurements$algorithm, measurements$map_resolution)
DunnTest(avg_mem ~ algorithm_map_resolution_interaction, measurements, method='BH')
```
