---
title: "Energy analysis - Circular arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'circular', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements$energy <- measurements$energy / 1000
measurements
```

# Data exploration
```{r}
measurements %>%
  group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(energy, type = "full")

measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(energy, type = "full")
```

```{r}
hist <- gghistogram(measurements, x = "energy", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('Energy (J)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "energy_circular_histogram.pdf")
```

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "energy", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('Energy (J)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "energy_circular_box.pdf")
```
# ANOVA

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(energy ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey")
ggdensity(res, fill="lightgrey")

# Create a QQ plot of residuals
ggqqplot(res)
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal. Trying some transformations (sqrt, log, reciprocal) also gives no normal result.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

30 out of 32 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "energy", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, most points are around the normal line.  The normality assumption is therefore met.

## 4. Homogeneity of variance

```{r}
levene_test(energy ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is lower than 0.05, hence we reject H0. Thus, we cannot assume homogeneity of variance  among all groups.

This is also clearly visible when plotting the data: one group has an extremely high variance (exceeding 4x the other variances).

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(energy, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
The assumptions for ANOVA are violated, hence we resort to PERMANOVA (using permutations).

## Running PERMANOVA
```{r}
set.seed(2021)
aovp.out <- aovp(energy ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

Proceeding for both levels of the linear update, to investigate the three-way interaction.

## Follow-up: Analyse low linear update

```{r}
low_linear_update_measurements <- measurements[measurements$linear_update == 0.2, ]
low_linear_update_measurements
```

Again, we test the important assumptions for ANOVA:

### 2. Normality of residuals
We fit a model based on the significant factors that are left (and their interactions) as parameters.

```{r}
model  <- lm(energy ~ algorithm*map_resolution*angular_update, data = low_linear_update_measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey") + custom_theme
ggdensity(res, fill="lightgrey") + custom_theme

# Create a QQ plot of residuals
ggqqplot(res) + custom_theme
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
```
Not normal. Trying some transformations (sqrt, log, reciprocal) also gives no normal result.

### 3. Normality within groups
```{r}
shapiro_tests <- low_linear_update_measurements %>% group_by(algorithm, map_resolution, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

15 out of 16 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(low_linear_update_measurements[low_linear_update_measurements$algorithm == 'cartographer', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(low_linear_update_measurements[low_linear_update_measurements$algorithm == 'gmapping', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(low_linear_update_measurements[low_linear_update_measurements$algorithm == 'hector', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(low_linear_update_measurements[low_linear_update_measurements$algorithm == 'karto', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, most points are around the normal line.  The normality assumption is therefore met.

### 4. Homogeneity of variance

```{r}
levene_test(energy ~ algorithm*map_resolution*angular_update*linear_update, data = low_linear_update_measurements)
```
The p-value is lower than 0.05, hence we reject H0. Thus, we cannot assume homogeneity of variance  among all groups.

This is also clearly visible when plotting the data: one group has an extremely high variance (exceeding 4x the other variances).

```{r}
agg_measurements <- low_linear_update_measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(energy, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd") + custom_theme
agg_measurements[order(agg_measurements$sd),]
```

The assumptions for ANOVA are violated, hence we resort to PERMANOVA (using permutations).

### Running PERMANOVA
```{r}
set.seed(2022)
aovp.out <- aovp(energy ~ algorithm*map_resolution*angular_update, data=low_linear_update_measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```
The results indicate a significant main effect of both the algorithm and the angular update factors.

### Post hoc tests
Since there are only two treatments for the angular update (high and low) we do not need to perform a post hoc test for this main effect.

To analyse the main effect of the algorithm, we first assess normality of the groups and compare their variance.

```{r}
gghistogram(low_linear_update_measurements, 'energy', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- low_linear_update_measurements %>% group_by(algorithm)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

```{r}
ggqqplot(low_linear_update_measurements, "energy", color="algorithm", ggtheme=custom_theme)
```

```{r}
levene_test(energy ~ algorithm, data = low_linear_update_measurements)
```
The assumptions for Tukey's HSB test are not met. Hence, we resort to the non-parametric Dunn test.

```{r}
DunnTest(energy ~ algorithm, low_linear_update_measurements, method='BH')
```

## Follow-up: Analyse high linear update

```{r}
high_linear_update_measurements <- measurements[measurements$linear_update == 1.0, ]
high_linear_update_measurements
```

Again, we test the important assumptions for ANOVA:

### 2. Normality of residuals
We fit a model based on the significant factors that are left (and their interactions) as parameters.

```{r}
model  <- lm(energy ~ algorithm*map_resolution*angular_update, data = high_linear_update_measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey") + custom_theme
ggdensity(res, fill="lightgrey") + custom_theme

# Create a QQ plot of residuals
ggqqplot(res) + custom_theme
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
```
The data seems normally distributed.

### 3. Normality within groups
```{r}
shapiro_tests <- high_linear_update_measurements %>% group_by(algorithm, map_resolution, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

15 out of 16 groups have a p-value above 0.05, so for these we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(high_linear_update_measurements[high_linear_update_measurements$algorithm == 'cartographer', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(high_linear_update_measurements[high_linear_update_measurements$algorithm == 'gmapping', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(high_linear_update_measurements[high_linear_update_measurements$algorithm == 'hector', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(high_linear_update_measurements[high_linear_update_measurements$algorithm == 'karto', ], "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, most points are around the normal line.  The normality assumption is therefore met.

### 4. Homogeneity of variance

```{r}
levene_test(energy ~ algorithm*map_resolution*angular_update*linear_update, data = high_linear_update_measurements)
```
The p-value is larger than 0.05, hence we fail to reject H0. Thus, we can assume homogeneity of variance among all groups.

### Running ANOVA
We can use the standard ANOVA, since all assumptions are met.
```{r}
aov.out <- aov(energy ~ algorithm*map_resolution*angular_update, data=high_linear_update_measurements)
summary(aov.out)
```
We observe a significant two-way interaction between the algorithm and angular update.

### Post hoc tests
To analyse the interaction and its components, we first assess normality of the groups and compare their variance.

```{r}
shapiro_tests <- high_linear_update_measurements %>% group_by(algorithm, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```
16 out of 16 groups have a p-value above 0.05, so we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(high_linear_update_measurements, "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ algorithm, labeller = "label_both")
```

```{r}
levene_test(energy ~ algorithm : angular_update, data = high_linear_update_measurements)
```
The assumptions for Tukey's HSB test are met.

```{r}
high_linear_update_measurements %>% tukey_hsd(energy ~ algorithm * angular_update)
```
