---
title: "Energy analysis - Point to point arena"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements <- measurements[measurements$arena == 'point_to_point', ]
measurements$algorithm <- as.factor(measurements$algorithm)
measurements$map_resolution <- as.factor(measurements$map_resolution)
measurements$linear_update <- as.factor(measurements$linear_update)
measurements$angular_update <- as.factor(measurements$angular_update)
measurements$energy <- measurements$energy / 1000
measurements
```

# Data exploration
Now, let us consider some descriptive statistics about the distribution of the measurements per group.

```{r}
measurements %>%
  group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(energy, type = "full")

measurements %>%
  group_by(algorithm) %>%
  get_summary_stats(energy, type = "full")
```

And a histogram of the distribution of the power measurements, which will be the primary focus of this analysis.

```{r}
hist <- gghistogram(measurements, x = "energy", color = "algorithm", fill = "algorithm", palette='jco') + custom_theme + xlab('Energy (J)') + ylab('Count') + labs(color='Algorithm', fill='Algorithm') 
hist
ggexport(hist, filename = "energy_point_to_point_histogram.pdf")
```
We notice the the gmapping measurements are very much spread out. The other measurements seem to be more clustered by algorithm.

To get an impression of the distribution of the distribution of average power per group, we plot the data in boxplots.

```{r}
bxp <- ggboxplot(measurements, x = "map_resolution", y = "energy", palette = "jco", color="algorithm", facet.by = c("linear_update", "angular_update"), labeller="label_both") + theme_bw(base_size = 19) + theme(legend.position="top", text=element_text(family="serif")) + xlab('Map resolution') + ylab('Energy (J)') + labs(color='Algorithm', fill='Algorithm')
bxp
ggexport(bxp, filename = "energy_point_to_point_box.pdf")
```
# ANOVA

## Assumption checking
ANOVA has four assumptions, which are checked in this section.

### 1. Independence of samples
By the design of the experiment, samples are independent (and fully randomised).

### 2. Normality of residuals
We fit a model based on all four factors (and their interactions) as parameters.
```{r}
model  <- lm(energy ~ algorithm*map_resolution*linear_update*angular_update, data = measurements)
res <- residuals(model)
```

Get an impression of the distribution of the residuals
```{r}
gghistogram(res, fill="lightgrey") + custom_theme
ggdensity(res, fill="lightgrey") + custom_theme

# Create a QQ plot of residuals
ggqqplot(res) + custom_theme
# Compute Shapiro-Wilk test of normality
shapiro_test(res)
skewness(res)
```
Not normal. Trying some transformations (sqrt, log, reciprocal) also gives no normal result.

### 3. Normality within groups
```{r}
shapiro_tests <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

32 out of 32 groups have a p-value above 0.05, so we cannot reject H0 (the data is normally distributed). Checking the qq plots in addition.

```{r}
ggqqplot(measurements[measurements$algorithm == 'cartographer', ], "energy", ggtheme = theme_bw()) + facet_grid( linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'gmapping', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'hector', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
ggqqplot(measurements[measurements$algorithm == 'karto', ], "energy", ggtheme = theme_bw()) + facet_grid(linear_update + angular_update ~ map_resolution, labeller = "label_both")
```

In the plots above, most points are around the normal line. The normality assumption is therefore met.

## 4. Homogeneity of variance

```{r}
levene_test(energy ~ algorithm*map_resolution*angular_update*linear_update, data = measurements)
```
The p-value is higher than 0.05, hence we fail to reject H0. Thus, we can assume homogeneity of variance  among all groups.

It is important to note that one group has an extremely high variance (exceeding 2x the other variances).

```{r}
agg_measurements <- measurements %>% group_by(algorithm, map_resolution, linear_update, angular_update) %>%
  get_summary_stats(energy, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```
The assumptions for ANOVA are violated, hence we resort to PERMANOVA (using permutations).

## Running PERMANOVA
```{r}
set.seed(2031)
aovp.out <- aovp(energy ~ algorithm*map_resolution*linear_update*angular_update, data=measurements, maxIter=100000000, nCycle=1)
summary(aovp.out)
```

The results indicate a significant main effect of the algorithm, linear update and angular update factors.

There is also a significant interaction between the algorithm and angular update.

### Post hoc tests
Since there are only two treatments for the linear update (high and low) we do not need to perform a post hoc test for this main effect.

To analyse the main effect of the algorithm and angular update (interaction), we first assess normality of the groups and compare their variance.

```{r}
gghistogram(measurements, 'energy', color="algorithm", fill="algorithm") + custom_theme

shapiro_tests <- measurements %>% group_by(algorithm, angular_update)  %>% shapiro_test(energy)
shapiro_tests[order(shapiro_tests$p),]
```

```{r}
ggqqplot(measurements, "energy", ggtheme = theme_bw()) + facet_grid(angular_update ~ algorithm, labeller = "label_both")
```

```{r}
levene_test(energy ~ algorithm*angular_update, data = measurements)
```

```{r}
agg_measurements <- measurements %>% group_by(algorithm, angular_update) %>%
  get_summary_stats(energy, type = "mean_sd")
ggscatter(agg_measurements, "mean", "sd")
agg_measurements[order(agg_measurements$sd),]
```

Both the normality of groups and heterogeneity of variance assumptions are violated, even after transformations.
We therefore use the non-parametric Dunn's test.

```{r}
algorithm_angular_update_interaction <- interaction(measurements$algorithm, measurements$angular_update)
DunnTest(energy ~ algorithm_angular_update_interaction, measurements, method='BH')
```

