---
title: "Result box computations"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lmPerm)
library(e1071)
library(DescTools)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
custom_theme <- theme_bw(base_size = 19) + theme(
  axis.line = element_line(colour = "black"),
  panel.border = element_blank(),
  legend.position = "top",
  text=element_text(family="serif"),
)
```

# Measurements
Let us first load all measurements obtained from the experiment.

```{r}
measurements <- read.csv("/home/engel/Documents/experiments/slam_experiment/run_table.csv")
measurements$avg_mem <- measurements$avg_mem / 1048576
measurements$energy <- measurements$energy / 1000

measurements
```

# Energy consumption
## Point to point arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'point_to_point'),]
```

### Simple main effect: linear update
```{r}
arena_measurements %>% group_by(linear_update) %>% get_summary_stats(energy, type = "mean_sd")
```

### Two-way interaction: algorithm and angular update
```{r}
arena_measurements %>% group_by(algorithm, angular_update) %>% get_summary_stats(energy, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(822.696, 826.567, 837.811))
group_2_mean <- mean(c(810.523, 807.320, 807.517, 809.296, 805.392))
group_1_mean
group_2_mean
group_1_mean / group_2_mean
```

## Circular arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'circular'),]
```

### Low linear update
```{r}
arena_measurements_low_linear <- measurements[(measurements$arena == 'circular') & (measurements$linear_update == 0.2), ]
```

```{r}
arena_measurements_low_linear %>% group_by(angular_update) %>% get_summary_stats(energy, type = "mean_sd")
```
```{r}
947.202 / 940.761
```


#### Algorithm groups
```{r}
arena_measurements_low_linear %>% group_by(algorithm) %>% get_summary_stats(energy, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(955.888, 951.649))
group_2_mean <- mean(c(934.840, 933.550))
group_1_mean
group_2_mean
group_1_mean / group_2_mean
```

### High linear update
```{r}
arena_measurements_high_linear <- measurements[(measurements$arena == 'circular') & (measurements$linear_update == 1.0), ]
```

```{r}
arena_measurements_high_linear %>% group_by(algorithm, angular_update) %>% get_summary_stats(energy, type = "mean_sd")
```
```{r}
973.357 / 936.823
```


#### Algorithm groups
```{r}
arena_measurements_high_linear %>% group_by(algorithm, angular_update) %>% get_summary_stats(energy, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(956.212, 953.481, 973.357))
group_2_mean <- mean(c(936.823, 933.671, 933.997, 933.624, 930.640))
group_1_mean
group_2_mean
group_1_mean / group_2_mean
```

# CPU utilisation
## Point to point arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'point_to_point'),]
```

### Simple main effect: map resolution
```{r}
arena_measurements %>% group_by(map_resolution) %>% get_summary_stats(avg_cpu, type = "mean_sd")
```

### Two-way interaction: algorithm and angular update
```{r}
arena_measurements %>% group_by(algorithm, angular_update) %>% get_summary_stats(avg_cpu, type = "mean_sd")
```

```{r}
41.982 / 31.651
group_1_mean <- mean(c(43.830, 42.373, 41.982))
group_2_mean <- mean(c(31.651, 28.657, 29.101, 29.454, 28.836))
group_1_mean
group_2_mean
group_1_mean - group_2_mean
group_1_mean / group_2_mean
```

## Circular arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'circular'),]
```

```{r}
arena_measurements %>% group_by(map_resolution) %>% get_summary_stats(avg_cpu, type = "mean_sd")
```

```{r}
35.173 - 34.316
35.173 / 34.316
```

### Algorithm groups
```{r}
arena_measurements %>% group_by(algorithm, angular_update) %>% get_summary_stats(avg_cpu, type = "mean_sd")
```

```{r}
44.511 - 31.636
44.511 / 31.636
group_1_mean <- mean(c(43.323, 42.593, 44.511))
group_2_mean <- mean(c(31.636, 28.585, 28.624, 29.866, 28.819))
group_1_mean
group_2_mean
group_1_mean - group_2_mean
group_1_mean / group_2_mean
```


# Memory utilisation
## Point to point arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'point_to_point'),]
```

### Two-way interaction: algorithm and map resolution
```{r}
arena_measurements %>% group_by(algorithm, map_resolution) %>% get_summary_stats(avg_mem, type = "mean_sd")
arena_measurements %>% group_by(algorithm) %>% get_summary_stats(avg_mem, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(624.663, 622.311, 623.170))
group_2_mean <- mean(c(636.445, 624.663))
group_3_mean <- mean(c(657.713, 657.653))
group_4_mean <- mean(c(744.417, 745.624))

group_1_mean
group_2_mean
group_3_mean
group_4_mean

group_2_mean / group_1_mean
group_3_mean / group_1_mean
group_4_mean / group_1_mean
```

## Circular arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'circular'),]
```

```{r}
arena_measurements %>% group_by(angular_update) %>% get_summary_stats(avg_mem, type = "mean_sd")
```

```{r}
666.025 - 663.909
666.025 / 663.909
```

### Algorithm groups
```{r}
arena_measurements %>% group_by(algorithm, map_resolution) %>% get_summary_stats(avg_mem, type = "mean_sd")
arena_measurements %>% group_by(algorithm) %>% get_summary_stats(avg_mem, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(622.463, 623.683, 626.667))
group_2_mean <- mean(c(631.628, 626.667))
group_3_mean <- mean(c(659.519, 658.776))
group_4_mean <- mean(c(746.627, 750.371))

group_1_mean
group_2_mean
group_3_mean
group_4_mean

group_2_mean / group_1_mean
group_3_mean / group_1_mean
group_4_mean / group_1_mean
```

# ROS mmessage count
## Point to point arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'point_to_point'),]
```

### Per algorithm
```{r}
arena_measurements %>% group_by(algorithm) %>% get_summary_stats(rosbag_msg_count, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(40155.97, 40579.65))
group_2_mean <- mean(c(41198.32))
group_3_mean <- mean(c(68170.48))

group_1_mean
group_2_mean
group_3_mean

group_2_mean / group_1_mean
group_3_mean / group_1_mean
```

## Circular arena
```{r}
arena_measurements <- measurements[(measurements$arena == 'circular'),]
```

### Per algorithm
```{r}
arena_measurements %>% group_by(algorithm) %>% get_summary_stats(rosbag_msg_count, type = "mean_sd")
```

```{r}
group_1_mean <- mean(c(46746.65, 46359.22))
group_2_mean <- mean(c(47430.38))
group_3_mean <- mean(c(79666.23))

group_1_mean
group_2_mean
group_3_mean

group_2_mean / group_1_mean
group_3_mean / group_1_mean
```
